---
title: "Burn Severity"
author: "Meagan Brown"
date: "2023-05-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Burn Severity

```{r}
#load libraries
library(tidyverse)
library(dplyr)
library(here)
library(janitor)
library(tmap)
library(stars)
library(gridExtra)
library(raster)
library(terra)
```

```{r}
# set working directory
setwd(here('~', 'meds', 'capstone', 'aquafire'))
```

### Read in Burn Severity All (1984 - 2021) .tif file

```{r}
burn_severity_stack <- rast(here::here('raster_output', 'burn_severity_all.tif'))
```

### Read in Burn Severity Reclassified

```{r}
burn_severity_recl <- rast(here::here('raster_output', 'burn_severity_recl.tif'))
```

### Read in Burn Severity Mode

```{r}
burn_severity_mode <- rast(here::here('raster_output', 'burn_severity_mode.tif'))
```

## Calculations

### Load necessary data
```{r}
gde <- read_sf(here::here('data', 'gde')) %>% 
  st_transform("EPSG: 3310")

# ecoregions data
eco_regions <- read_sf(here::here('data', 'ca_eco_l3')) %>% 
  janitor::clean_names()  %>% 
  st_transform("EPSG: 3310") %>% 
  rename(region = us_l3name)
```

# mode value in GDEs vs non-GDEs
```{r}
# Filter for just the Southern California/Northern Baja Coast Ecoregion
socal_baja_coast <- eco_regions %>% 
  filter(region == 'Southern California/Northern Baja Coast')

# crop burn severity and ecoregion
p <- crop(burn_severity_mode, socal_baja_coast)

tmap_mode('view')

tm_shape(p) +
    tm_raster(style="cat", palette="-viridis", title="Discrete Raster") +
    tm_layout(legend.outside = TRUE, legend.outside.position = "right")
```

## Creating Files if you don't already have it
### Burn Severity All (1984 - 2021) for loop - takes 3 hours to load

```{r}
# Load the necessary library
library(terra)

# Define the directory where the raster files are located
raster_dir <- here::here("data", "MTBS_BSmosaics")

# Define the target CRS
target_crs <- "EPSG:3310"

# Get a list of .tif files in the directory
tif_files <- list.files(raster_dir, pattern = "\\.tif$", full.names = TRUE)

# Initialize the raster stack variable
raster_stack <- NULL

# Loop through the files
for (file_path in tif_files) {
  # Print a message to the console
  message("Processing file: ", file_path)

  # Read the raster file
  raster <- rast(file_path)
  
  # Project the raster to the target CRS
  raster_new_crs <- project(raster, target_crs)
  
  # If this is the first file, initialize the raster stack
  if (is.null(raster_stack)) {
    raster_stack <- raster_new_crs
  } else {
    # Otherwise, resample the raster to match the first, and add it to the stack
    raster_resampled <- resample(raster_new_crs, raster_stack, method = "near")
    raster_stack <- c(raster_stack, raster_resampled)
  }
}

writeRaster(raster_stack, file.path(here::here('raster_output', 'burn_severity_all.tif')))

# Plot the raster stack
plot(raster_stack)
```

### Reclassify Burn Severity Values Across Raster Stack - takes 3 hours to load

```{r}
# Value	Legend
# 0	Unburnt
# 1	unburnt/low

# 2	low
# 3	moderate
# 4	high

# 5	increased greenness
# 6	masked

# Define reclassification matrix
bs_recl <- c(-Inf, 1, NA,   # Unburnt
             1, 2, NA,      # unburnt/low
             2, 3, 2,       # low
             3, 4, 3,       # moderate
             4, 5, 4,       # high
             5, 6, NA,      # increased greenness
             6, Inf, NA)    # masked

# Convert to a matrix with 3 columns
bs_matrix <- matrix(bs_recl, ncol = 3, byrow = TRUE)

# Apply reclassification to the raster stack
burn_severity_recl <- classify(burn_severity_stack, rcl = bs_matrix)

burn_severity_recl

# writeRaster(burn_severity_recl, file.path(here::here('raster_output', 'burn_severity_recl.tif')))
```

### Burn Severity Mode Across Raster Stack - takes 3 hours to load

```{r}
# calculate the mode value for each cell across layers
mode_value <- function(x) {
  ux <- unique(x)    
  ux = ux[!is.na(ux)]
  mode = ux[which.max(tabulate(match(x, ux)))]
  mode
}

burn_severity_mode = terra::app(burn_severity_recl, fun = mode_value)

# writeRaster(burn_severity_mode, file.path(here::here('raster_output', 'burn_severity_mode.tif')), overwrite = TRUE)

plot(burn_severity_mode)
```


