---
title: "Burn Severity Stats"
author: "Jillian Allison"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Run the Burn Severity and GDE sections of the Objects.RMD file to get all spatial objects necessary to run this code. Download necessary files and put them in the raster_output folder before running. 

### Eco Region Key: 

   **code**                                           **region**
          1                                         Coast Range
         13                             Central Basin and Range
         14                              Mojave Basin and Range
          4                                            Cascades
          5                                       Sierra Nevada
          6  Central California Foothills and Coastal Mountains
          7                           Central California Valley
         78 Klamath Mountains/California High North Coast Range
          8                       Southern California Mountains
         80                            Northern Basin and Range
         81                             Sonoran Basin and Range
         85             Southern California/Northern Baja Coast
          9               Eastern Cascades Slopes and Foothills
          
```{r}
library(tidyverse)
library(sf)
library(here)
library(janitor)
library(raster)
library(rgdal)
library(fasterize)
library(exactextractr)
library(dplyr)
library(gridExtra)
library(kableExtra)
source("user/jillian/functions/remove_ag_urban.R")
source("user/jillian/functions/raster_to_ecoregions.R")
source("user/jillian/functions/stratified_sample.R")
```

# Sample 
Many areas have no values for burn severity, so I'll mask the GDE raster layer to only have data for areas that have values for burn severity. 


```{r}
set.seed(123) 
burn_severity_1_sample <- stratified_sample(burn_severity_1, gde_1, n = 1000)
burn_severity_4_sample <- stratified_sample(burn_severity_4, gde_4, n = 1000)
burn_severity_5_sample <- stratified_sample(burn_severity_5, gde_5, n = 1000)
burn_severity_6_sample <- stratified_sample(burn_severity_6, gde_6, n = 1000)
burn_severity_7_sample <- stratified_sample(burn_severity_7, gde_7, n = 1000)
burn_severity_8_sample <- stratified_sample(burn_severity_8, gde_8, n = 1000)
burn_severity_9_sample <- stratified_sample(burn_severity_9, gde_9, n = 1000)
burn_severity_13_sample <- stratified_sample(burn_severity_13, gde_13, n = 1000)
burn_severity_14_sample <- stratified_sample(burn_severity_14, gde_14, n = 1000)
burn_severity_78_sample <- stratified_sample(burn_severity_78, gde_78, n = 1000)
burn_severity_80_sample <- stratified_sample(burn_severity_80, gde_80, n = 1000)
burn_severity_81_sample <- stratified_sample(burn_severity_81, gde_81, n = 1000)
burn_severity_85_sample <- stratified_sample(burn_severity_85, gde_85, n = 1000)
```

# Summary Statistics 
```{r}

```

# Basic Visualizations


# T-Test 
