---
title: "Fire Count Stats"
author: "Jillian Allison"
date: "`r Sys.Date()`"
output: html_document
---

Run the Fire Count and GDE sections of the Objects.RMD file to get all spatial objects necessary to run this code. Download necessary files and put them in the raster_output folder before running. 

### Eco Region Key: 

   **code**                                           **region**
          1                                         Coast Range
         13                             Central Basin and Range
         14                              Mojave Basin and Range
          4                                            Cascades
          5                                       Sierra Nevada
          6  Central California Foothills and Coastal Mountains
          7                           Central California Valley
         78 Klamath Mountains/California High North Coast Range
          8                       Southern California Mountains
         80                            Northern Basin and Range
         81                             Sonoran Basin and Range
         85             Southern California/Northern Baja Coast
          9               Eastern Cascades Slopes and Foothills

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(sf)
library(boot)
library(boot.pval)
library(here)
library(janitor)
library(raster)
library(tmap)
library(rgdal)
library(fasterize)
library(exactextractr)
library(dplyr)
library(gridExtra)
library(kableExtra)
source("user/jillian/functions/remove_ag_urban.R")
source("user/jillian/functions/raster_to_ecoregions.R")
source("user/jillian/functions/stratified_sample.R")
source("user/jillian/functions/fire_count_violin_plot.R")
source("user/jillian/functions/fire_count_stats.R")
source("user/jillian/functions/diff_in_means_bootstrap.R")
source("user/jillian/functions/bootstrap_results.R")
source("user/jillian/functions/fire_count_histogram.R")
```

# Necessary Objects 
```{r}
# For eco region names / codes 
crs_ca <- st_crs(3310)
# Load in ecoregions shapefile
eco_regions <- read_sf(here::here('data', 'ca_eco_l3')) %>% 
  janitor::clean_names()  %>% 
  st_transform(crs_ca) %>% 
  rename(region = us_l3name)  %>%
  mutate(code = as.numeric(us_l3code))
# Create file containing eco region names 
eco_regions_names <- as.data.frame(eco_regions) %>%
  subset(select = c("us_l3code", "region"))
eco_regions_geoms <- as.data.frame(eco_regions) %>%
  subset(select = c("us_l3code", "geometry"))

gde_1 <- raster("raster_output/gde_1.tif")
gde_4 <- raster("raster_output/gde_4.tif")
gde_5 <- raster("raster_output/gde_5.tif")
gde_6 <- raster("raster_output/gde_6.tif")
gde_7 <- raster("raster_output/gde_7.tif")
gde_8 <- raster("raster_output/gde_8.tif")
gde_9 <- raster("raster_output/gde_9.tif")
gde_13 <- raster("raster_output/gde_13.tif")
gde_14 <- raster("raster_output/gde_14.tif")
gde_78 <- raster("raster_output/gde_78.tif")
gde_80 <- raster("raster_output/gde_80.tif")
gde_81 <- raster("raster_output/gde_81.tif")
gde_85 <- raster("raster_output/gde_85.tif")

fire_count_1 <- raster("raster_output/fire_count_1.tif")
fire_count_4 <- raster("raster_output/fire_count_4.tif")
fire_count_5 <- raster("raster_output/fire_count_5.tif")
fire_count_6 <- raster("raster_output/fire_count_6.tif")
fire_count_7 <- raster("raster_output/fire_count_7.tif")
fire_count_8 <- raster("raster_output/fire_count_8.tif")
fire_count_9 <- raster("raster_output/fire_count_9.tif")
fire_count_13 <- raster("raster_output/fire_count_13.tif")
fire_count_14 <- raster("raster_output/fire_count_14.tif")
fire_count_78 <- raster("raster_output/fire_count_78.tif")
fire_count_80 <- raster("raster_output/fire_count_80.tif")
fire_count_81 <- raster("raster_output/fire_count_81.tif")
fire_count_85 <- raster("raster_output/fire_count_85.tif")
```


# Samples 
```{r}
set.seed(123) 
fire_count_1_sample <- stratified_sample(fire_count_1, gde_1, n = 1000)
fire_count_4_sample <- stratified_sample(fire_count_4, gde_4, n = 1000)
fire_count_5_sample <- stratified_sample(fire_count_5, gde_5, n = 1000)
fire_count_6_sample <- stratified_sample(fire_count_6, gde_6, n = 1000)
fire_count_7_sample <- stratified_sample(fire_count_7, gde_7, n = 1000)
fire_count_8_sample <- stratified_sample(fire_count_8, gde_8, n = 1000)
fire_count_9_sample <- stratified_sample(fire_count_9, gde_9, n = 1000)
fire_count_13_sample <- stratified_sample(fire_count_13, gde_13, n = 1000)
fire_count_14_sample <- stratified_sample(fire_count_14, gde_14, n = 1000)
fire_count_78_sample <- stratified_sample(fire_count_78, gde_78, n = 1000)
fire_count_80_sample <- stratified_sample(fire_count_80, gde_80, n = 1000)
fire_count_81_sample <- stratified_sample(fire_count_81, gde_81, n = 1000)
fire_count_85_sample <- stratified_sample(fire_count_85, gde_85, n = 1000)

# Show locations of sample points 
ggplot(fire_count_5_sample, aes(x = x, y = y)) + 
  geom_point() + 
  theme_minimal()
```

# Bootstrapping 
```{r}
eco_region_codes <- c(1, 13, 14, 4, 5, 6, 7, 78, 8, 80, 81, 85, 9)
fire_count_sample_list <- list()

for(i in 1:13){
  obj <- (paste0('fire_count_', eco_region_codes[i], '_sample'))
  fire_count_sample_list <- c(fire_count_sample_list, obj)
}

fire_count_bootstrap <- bootstrap_results(fire_count_sample_list) 

fire_count_results <- fire_count_bootstrap %>%
  bind_cols(eco_regions_names) %>%
  inner_join(eco_regions_geoms, by = "us_l3code") %>% 
  mutate(result = case_when(
    p_value < 0.05 & confidence_interval_upper < 0 ~ "Reject H0 - Incr. Fire Count",
    p_value < 0.05 & confidence_interval_lower > 0 ~ "Reject H0 - Decr. Fire Count",
    p_value > 0.05 ~ "Cannot Reject H0"))

fire_count_results_sp <- fire_count_results %>%
  subset(select = c("p_value","result","region", "geometry")) %>%
  st_as_sf(crs = crs_ca$wkt) 

```

# Map showing results from difference in mean test
```{r}
# grey is #EDF0F0
# red is #9D2A38
# teal is #2A9D8F
# cant reject is #C8D5D3
color_palette <- c("Reject H0 - Incr. Fire Count" = "#9D2A38",
                   "Reject H0 - Decr. Fire Count" = "#2A9D8F",
                   "Cannot Reject H0" = "#C8D5D3")

color_palette_presentation <- c("Coast Range" = "#EDF0F0",
                                "Cascades" = "#EDF0F0",
                                "Sierra Nevada" = "#EDF0F0", 
                                "Central California Foothills and Coastal Mountains" = "#EDF0F0",
                                "Klamath Mountains/California High North Coast Range" = "#EDF0F0", 
                                "Southern California Mountains" = "#2A9D8F",
                                "Northern Basin and Range" = "#EDF0F0", 
                                "Sonoran Basin and Range" = "#EDF0F0", 
                                "Southern California/Northern Baja Coast" = "#EDF0F0",
                                "Eastern Cascades Slopes and Foothills" = "#EDF0F0",
                                "Central Basin and Range" = "#EDF0F0", 
                                "Mojave Basin and Range" = "#EDF0F0", 
                                "Central California Valley" = "#EDF0F0")

tm_shape(fire_count_results_sp) + 
  tm_polygons + 
  tm_fill(col = "result", palette = color_palette, title = "Result") + 
  tm_borders(col = "white", 
             lwd = 0.5) + 
  tm_layout(main.title = "",
            main.title.size = 0.8,
            frame = FALSE) +  
  tm_legend(position = c(1, -1))
```

# Summary Stats
```{r}
fire_count_1_stats <- fire_count_stats(fire_count_1_sample, eco_region_code = 1)
fire_count_4_stats <- fire_count_stats(fire_count_4_sample, eco_region_code = 4)
fire_count_5_stats <- fire_count_stats(fire_count_5_sample, eco_region_code = 5)
fire_count_6_stats <- fire_count_stats(fire_count_6_sample, eco_region_code = 6)
fire_count_7_stats <- fire_count_stats(fire_count_7_sample, eco_region_code = 7)
fire_count_8_stats <- fire_count_stats(fire_count_8_sample, eco_region_code = 8)
fire_count_9_stats <- fire_count_stats(fire_count_9_sample, eco_region_code = 9)
fire_count_13_stats <- fire_count_stats(fire_count_13_sample, eco_region_code = 13)
fire_count_14_stats <- fire_count_stats(fire_count_14_sample, eco_region_code = 14)
fire_count_78_stats <- fire_count_stats(fire_count_78_sample, eco_region_code = 78)
fire_count_80_stats <- fire_count_stats(fire_count_80_sample, eco_region_code = 80)
fire_count_81_stats <- fire_count_stats(fire_count_81_sample, eco_region_code = 81)
fire_count_85_stats <- fire_count_stats(fire_count_85_sample, eco_region_code = 85)

fire_count_stats_all <- rbind(fire_count_1_stats, fire_count_4_stats, 
                              fire_count_5_stats, fire_count_6_stats,
                              fire_count_7_stats, fire_count_8_stats,
                              fire_count_9_stats, fire_count_13_stats, 
                              fire_count_14_stats, fire_count_78_stats,
                              fire_count_80_stats, fire_count_81_stats,
                              fire_count_85_stats)
```

# Basic Visualizations 


### Histograms
```{r}
fire_count_sample_list
eco_region_codes


# # Loop through each value in the list and eco region code vector
# for (i in seq_along(fire_count_sample_list)) {
#   # Get the current object name and eco region code
#   sample <- get(fire_count_sample_list[[i]])
#   eco_region_code <- eco_region_codes[i]
#   
#   # Call the function with the object name and eco region code
#   plot <- fire_count_histogram(sample, eco_region_code)
#   
#   # Save the plot as a PNG file
#   png_file <- paste0("plots/fire_count_histogram_", eco_region_code, ".png")
#   png(png_file, width = 1024, height = 600, res = 200)
#   
#   ggsave(
#     filename = paste0("fire_count_histogram_", eco_region_code, ".png"), 
#     plot = plot, 
#     device = "png", 
#     path = "plots/", 
#     units = c("in"),
#     width = 11,
#     height = 7
#   )
# dev.off()
# }

fire_count_shiny_histogram_df <- data.frame() 

# Loop through each value in the list and eco region code vector
for (i in seq_along(fire_count_sample_list)) {
  # Get the current object name and eco region code
  sample <- get(fire_count_sample_list[[i]])
  eco_region_code <- eco_region_codes[i]
  
  df <- fire_count_histogram(sample, eco_region_code)
  
  fire_count_shiny_histogram_df <- rbind(fire_count_shiny_histogram_df, df)
  
}

```

### Violin Plots
```{r fig.align="center", echo = FALSE,fig.width = 14}
plot_fc_1 <- fire_count_violin_plot(fire_count_1_sample, eco_region_code = 1)
plot_fc_4 <- fire_count_violin_plot(fire_count_4_sample, eco_region_code = 4)
plot_fc_5 <- fire_count_violin_plot(fire_count_5_sample, eco_region_code = 5)
plot_fc_6 <- fire_count_violin_plot(fire_count_6_sample, eco_region_code = 6)
plot_fc_7 <- fire_count_violin_plot(fire_count_7_sample, eco_region_code = 7)
plot_fc_8 <- fire_count_violin_plot(fire_count_8_sample, eco_region_code = 8)
plot_fc_9 <- fire_count_violin_plot(fire_count_9_sample, eco_region_code = 9)
plot_fc_13 <- fire_count_violin_plot(fire_count_13_sample, eco_region_code = 13)
plot_fc_14 <- fire_count_violin_plot(fire_count_14_sample, eco_region_code = 14)
plot_fc_78 <- fire_count_violin_plot(fire_count_78_sample, eco_region_code = 78)
plot_fc_80 <- fire_count_violin_plot(fire_count_80_sample, eco_region_code = 80)
plot_fc_81 <- fire_count_violin_plot(fire_count_81_sample, eco_region_code = 81)
plot_fc_85 <- fire_count_violin_plot(fire_count_85_sample, eco_region_code = 85)

plot_fc_5
plot_fc_1
plot_fc_85
plot_fc_8 

plot_fc_5
```

### Original Histogram Code: Histogram Comparing GDE Dist and Non-GDE Dist (now a function)
```{r}
## CHECK VARIABLE NAMES IM JUST TESTING DIFFERENT ECO REGIONS WITH THIS 
fire_count_histogram(sample_data = fire_count_85_sample, eco_region_code = 85)

fire_count_gdes <- fire_count_85_sample %>%
  filter(gde == 1) %>%
  group_by(value) %>%
  summarize(proportion_1 = n() / nrow(.) * 100)

fire_count_nongdes <- fire_count_5_sample %>%
  filter(gde == 0) %>%
  group_by(value) %>%
  summarize(proportion_0 = n() / nrow(.) * 100)

fire_count_histogram_obj <- inner_join(fire_count_gdes, fire_count_nongdes, by = "value") %>%
  rename("GDE" = "proportion_1", 
         "NonGDE" = "proportion_0") %>% 
  pivot_longer(cols = c(2,3),
               names_to = "gde_status",
               values_to = "proportion"
               )

ggplot(fire_count_histogram_obj, aes(x = value, y = proportion, fill = gde_status)) + 
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("#A3B18A", "#DDA15E")) +
  labs(x = "Fire Count", 
       y = "Relative Frequency (%)",
       fill = "GDE Status",
       title = "Southern CA / Northern Baja Coast: Relative Frequency of Fire Counts in GDEs and Non-GDEs") + 
  theme_classic() + 
  theme(legend.position = 'none',
          plot.title = element_text(hjust = 0.5,
                                    size = 15),
          axis.text = element_text(size = 13,
                                   color = 'black'),
          axis.title = element_text(size = 15,
                                    color = 'black'),
          axis.title.x = element_text(vjust = -1.1),
          axis.text.x = element_text(vjust = -1.5)) + 
  scale_y_continuous(expand = c(0,0)) + scale_x_continuous(expand = c(0,0))

```





# Old Logistic Regressions

```{r}
# Coast Range
fire_count_1_sample$gde <- factor(fire_count_1_sample$gde)
fire_count_1_glm <- glm(value ~ gde, data = fire_count_1_sample, family = "poisson")

# Cascades 
fire_count_4_sample$gde <- factor(fire_count_4_sample$gde)
fire_count_4_glm <- glm(value ~ gde, data = fire_count_4_sample, family = "poisson")

# Sierra Nevada 
fire_count_5_sample$gde <- factor(fire_count_5_sample$gde)
fire_count_5_glm <- glm(value ~ gde, data = fire_count_5_sample, family = "poisson")

# Central CA Foothills and Coastal Mountains 
fire_count_6_sample$gde <- factor(fire_count_6_sample$gde)
fire_count_6_glm <- glm(value ~ gde, data = fire_count_6_sample, family = "poisson")

# Central California Valley
fire_count_7_sample$gde <- factor(fire_count_7_sample$gde)
fire_count_7_glm <- glm(value ~ gde, data = fire_count_7_sample, family = "poisson")

# Southern CA Mountains
fire_count_8_sample$gde <- factor(fire_count_8_sample$gde)
fire_count_8_glm <- glm(value ~ gde, data = fire_count_8_sample, family = "poisson")

# Eastern Cascades Slopes and Foothills *** 
fire_count_9_sample$gde <- factor(fire_count_9_sample$gde)
fire_count_9_glm <- glm(value ~ gde, data = fire_count_9_sample, family = "poisson")

# Central Basin and Range *** 
fire_count_13_sample$gde <- factor(fire_count_13_sample$gde)
fire_count_13_glm <- glm(value ~ gde, data = fire_count_13_sample, family = "poisson")

# Mojave Basin and Range * 
fire_count_14_sample$gde <- factor(fire_count_14_sample$gde)
fire_count_14_glm <- glm(value ~ gde, data = fire_count_14_sample, family = "poisson")

# Klamath Mountains/California High North Coast Range *** 
fire_count_78_sample$gde <- factor(fire_count_78_sample$gde)
fire_count_78_glm <- glm(value ~ gde, data = fire_count_78_sample, family = "poisson")

# Northern Basin and Range *** 
fire_count_80_sample$gde <- factor(fire_count_80_sample$gde)
fire_count_80_glm <- glm(value ~ gde, data = fire_count_80_sample, family = "poisson")

# Sonoran Basin and Range *
fire_count_81_sample$gde <- factor(fire_count_81_sample$gde)
fire_count_81_glm <- glm(value ~ gde, data = fire_count_81_sample, family = "poisson")

# Southern CA and Northern Baja Coast *** 
fire_count_85_sample$gde <- factor(fire_count_85_sample$gde)
fire_count_85_glm <- glm(value ~ gde, data = fire_count_85_sample, family = "poisson")
```



# Old Summary tables 
```{r}
pretty_6 <- kable(fire_count_6_stats %>% select(-ecoregion_code), align = "c", col.names = c("Group", "Mean Fire Count", "Maximum Fire Count", "Mode (not including zeros)"), caption = "Central California Foothills and Coastal Mountains") %>%
  kable_styling(bootstrap_options = "striped")

pretty_85 <- kable(fire_count_85_stats %>% select(-ecoregion_code), align = "c", col.names = c("Group", "Mean Fire Count", "Maximum Fire Count", "Mode (not including zeros)"), caption = "Southern California/Northern Baja Coast") %>%
  kable_styling(bootstrap_options = "striped")

pretty_8 <- kable(fire_count_8_stats %>% select(-ecoregion_code), align = "c", col.names = c("Group", "Mean Fire Count", "Maximum Fire Count", "Mode (not including zeros)"), caption = "Southern California Mountains") %>%
  kable_styling(bootstrap_options = "striped")

pretty_6
pretty_8
pretty_85
```


# Old violin plot code

```{r}
# Coast Range 
fire_count_1_sample <- stratified_sample(fire_count_1, gde_1, n = 1000)
fire_count_1_gdes <- fire_count_1_sample %>%
  filter(gde == 1)
fire_count_1_nongdes <- fire_count_1_sample %>%
  filter(gde == 0)

fire_count_1_stats <- fire_count_stats(fire_count_1_sample)


# Plot density curves for GDEs and non-GDEs
library(ggplot2)
# Create violin plot
ggplot(fire_count_1_sample, aes(x = gde, y = value, fill = gde)) +
  geom_violin(trim = FALSE) +
  scale_fill_manual(values = c("springgreen4", "dodgerblue"), name = "GDE Status") + 
  xlab("GDE Status") +
  ylab("Fire Count") +
  ggtitle("Distribution of Fire Counts in Non-GDEs vs GDEs") + 
  theme_minimal()


ggplot(fire_count_1_sample, aes(x = value, color = gde)) + 
  stat_ecdf() +
  scale_color_manual(values = c("springgreen4", "dodgerblue"), name = "GDE Status") +
  xlab("Fire Count") +
  ylab("Cumulative Proportion") +
  ggtitle("Cumulative Distribution of Fire Counts in GDEs vs Non-GDEs") + 
  theme_minimal()

```

```{r}
# Northern basin and range
fire_count_80_sample <- stratified_sample(fire_count_80, gde_80, n = 1000)

fire_count_80_stats <- fire_count_stats(fire_count_80_sample)

# Create violin plot
ggplot(fire_count_80_sample, aes(x = gde, y = value, fill = gde)) +
  geom_violin(trim = FALSE) +
  scale_fill_manual(values = c("#DDA15E", "#A3B18A")) +
  scale_x_discrete(labels = c('Non-GDE', 'GDE')) +
  xlab("") +
  ylab("Fire Count") +
  ggtitle("Distribution of Fire Counts in Non-GDEs vs GDEs") +
  theme_classic() +
  coord_flip() +
  theme(legend.position = 'none',
        plot.title = element_text(hjust = 0.5,
                                  size = 15),
        axis.text = element_text(size = 13,
                                 color = 'black'),
        axis.title = element_text(size = 15,
                                  color = 'black'),
        axis.title.x = element_text(vjust = -1.1),
        axis.text.x = element_text(vjust = -1.5))


  
```

```{r}
gde_6 <- raster("raster_output/gde_6.tif")
fire_count_6 <- raster("raster_output/fire_count_6.tif")
source("user/jillian/functions/stratified_sample.R")
source("user/jillian/functions/fire_count_stats.R")

set.seed(123)

# Central CA Foothills and Coastal Mountains
fire_count_6_sample <- stratified_sample(fire_count_6, gde_6, n = 1000)

fire_count_6_stats <- fire_count_stats(fire_count_6_sample)

# Create violin plot
ggplot(fire_count_6_sample, aes(x = gde, y = value, fill = gde)) +
  geom_violin(trim = FALSE) +
  scale_fill_manual(values = c("#DDA15E", "#A3B18A")) +
  scale_x_discrete(labels = c('Non-GDE', 'GDE')) +
  xlab("") +
  ylab("Fire Count") +
  ggtitle("Distribution of Fire Counts in Non-GDEs vs GDEs") +
  theme_classic() +
  coord_flip() +
  theme(legend.position = 'none',
        plot.title = element_text(hjust = 0.5,
                                  size = 15),
        axis.text = element_text(size = 13,
                                 color = 'black'),
        axis.title = element_text(size = 15,
                                  color = 'black'),
        axis.title.x = element_text(vjust = -1.1),
        axis.text.x = element_text(vjust = -1.5))

ggplot(fire_count_6_sample, aes(x = value, color = gde)) + 
  stat_ecdf() +
  scale_color_manual(values = c("springgreen4", "dodgerblue"), name = "GDE Status") +
  xlab("Fire Count") +
  ylab("Cumulative Proportion") +
  ggtitle("Cumulative Distribution of Fire Counts in GDEs vs Non-GDEs") + 
  theme_minimal()
  
```

# Original Bootstrapping Code
```{r}
# Step 1: Load the required package
library(boot)
library(boot.pval)

# Step 2: Define the function that calculates the test statistic (difference in means)
diff_in_means <- function(sample_data, indices) {
  gde_0 <- sample_data %>%
    filter(gde == 0)
  gde_1 <- sample_data %>%
    filter(gde == 1)
  
  nongde_sample <- gde_0[indices, 3]
  gde_sample <- gde_1[indices, 3]
  return(mean(nongde_sample, na.rm = TRUE) - mean(gde_sample, na.rm = TRUE))
}

# Step 4: Perform the bootstrap procedure
bootstrap_result <- boot(fire_count_85_sample, diff_in_means, R = 1000)

# Step 5: Obtain the bootstrap confidence interval and p-value
confidence_interval <- boot.ci(bootstrap_result, conf = 0.95, type = "norm")$normal
p_value <- boot.pval(bootstrap_result, theta_null = 0)

# Step 6: Print the results
print("Bootstrap Confidence Interval:")
print(confidence_interval)
print("P-value:")
print(p_value)

diff_in_means_bootstrap(fire_count_85_sample)


```

```{r}
# fire_count_sample <- stratified_sample(fire_count_raster, gde_raster, n = 1000)

# write.csv(fire_count_sample, "user/jillian/csv213/fire_count_sample.csv", row.names=FALSE)
```

- comparing distributions w two plots like two histograms 
- t test - chi sq 
- anova comparing eco-regions for all of CA analysis 
- visualization on map of CA maybe w like colors showing quantified relationship 
- write up about it 