---
title: "Fire Severity Stats"
author: "Jillian Allison"
date: "`r Sys.Date()`"
output: html_document
---

### Eco Region Key: 

   **code**                                           **region**
          1                                         Coast Range
         13                             Central Basin and Range
         14                              Mojave Basin and Range
          4                                            Cascades
          5                                       Sierra Nevada
          6  Central California Foothills and Coastal Mountains
          7                           Central California Valley
         78 Klamath Mountains/California High North Coast Range
          8                       Southern California Mountains
         80                            Northern Basin and Range
         81                             Sonoran Basin and Range
         85             Southern California/Northern Baja Coast
          9               Eastern Cascades Slopes and Foothills

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(sf)
library(boot)
library(boot.pval)
library(here)
library(janitor)
library(raster)
library(tmap)
library(rgdal)
library(fasterize)
library(exactextractr)
library(dplyr)
library(gridExtra)
library(kableExtra)
source("user/jillian/functions/remove_ag_urban.R")
source("user/jillian/functions/raster_to_ecoregions.R")
source("user/jillian/functions/stratified_sample.R")
source("user/jillian/functions/fire_count_violin_plot.R")
source("user/jillian/functions/fire_count_stats.R")
source("user/jillian/functions/diff_in_means_bootstrap.R")
source("user/jillian/functions/bootstrap_results.R")
source("user/jillian/functions/fire_count_histogram.R")
```

## Burn Severity
```{r}
burn_severity_1 <- raster("raster_output/burn_severity_1.tif")
burn_severity_4 <- raster("raster_output/burn_severity_4.tif")
burn_severity_5 <- raster("raster_output/burn_severity_5.tif")
burn_severity_6 <- raster("raster_output/burn_severity_6.tif")
burn_severity_7 <- raster("raster_output/burn_severity_7.tif")
burn_severity_8 <- raster("raster_output/burn_severity_8.tif")
burn_severity_9 <- raster("raster_output/burn_severity_9.tif")
burn_severity_13 <- raster("raster_output/burn_severity_13.tif")
burn_severity_14 <- raster("raster_output/burn_severity_14.tif")
burn_severity_78 <- raster("raster_output/burn_severity_78.tif")
burn_severity_80 <- raster("raster_output/burn_severity_80.tif")
burn_severity_81 <- raster("raster_output/burn_severity_81.tif")
burn_severity_85 <- raster("raster_output/burn_severity_85.tif")
```

## GDEs
```{r}
gde_1 <- raster("raster_output/gde_1.tif")
gde_4 <- raster("raster_output/gde_4.tif")
gde_5 <- raster("raster_output/gde_5.tif")
gde_6 <- raster("raster_output/gde_6.tif")
gde_7 <- raster("raster_output/gde_7.tif")
gde_8 <- raster("raster_output/gde_8.tif")
gde_9 <- raster("raster_output/gde_9.tif")
gde_13 <- raster("raster_output/gde_13.tif")
gde_14 <- raster("raster_output/gde_14.tif")
gde_78 <- raster("raster_output/gde_78.tif")
gde_80 <- raster("raster_output/gde_80.tif")
gde_81 <- raster("raster_output/gde_81.tif")
gde_85 <- raster("raster_output/gde_85.tif")
```

## Mask GDEs to only areas where we have burn severity data
(For areas where there is an existing fire)
```{r}
# gde_85 - Socal Northern Baja Coast 
gde_85_masked <- mask(gde_85, burn_severity_85)

# gde_5 - Sierra Nevada 
gde_5_masked <- mask(gde_5, burn_severity_5)
```

## Sample 
```{r}
burn_sev_sampling <- function(input_raster_layer, gde_raster_layer, n = 30){
  
  gde_raster_layer <- mask(gde_raster_layer, input_raster_layer)
  
  gde_sample <- sampleStratified(gde_raster_layer, size = 1000, sp = TRUE, buffer = 100)
  
  init_raster_sample <- extract(input_raster_layer, gde_sample) 
  
  gde_sample$value <- init_raster_sample
  
  df <- as.data.frame(gde_sample) %>%
    mutate(gde = as.factor(layer)) %>%
    dplyr::select(cell, gde, value, x, y) 

  
  burn_sev_gde_sample <- df %>%
    filter(gde == 1) %>%
    sample_n(n)
  
  burn_sev_nongde_sample <- df %>%
    filter(gde == 0) %>%
    sample_n(n)
  
  burn_severity_sample <- rbind(burn_sev_gde_sample, burn_sev_nongde_sample)
  
  
  return(burn_severity_sample)
}
```

```{r}
set.seed(123) 

burn_severity_1_sample <- burn_sev_sampling(burn_severity_1, gde_1, n = 20) # n is the most samples I could take given the buffer- small area
burn_severity_4_sample <- burn_sev_sampling(burn_severity_4, gde_4)
burn_severity_5_sample <- burn_sev_sampling(burn_severity_5, gde_5)
burn_severity_6_sample <- burn_sev_sampling(burn_severity_6, gde_6)
# burn_severity_7_sample <- burn_sev_sampling(burn_severity_7, gde_7) not enough data 
burn_severity_8_sample <- burn_sev_sampling(burn_severity_8, gde_8)
burn_severity_9_sample <- burn_sev_sampling(burn_severity_9, gde_9)
# burn_severity_13_sample <- burn_sev_sampling(burn_severity_13, gde_13) not enough data
# burn_severity_14_sample <- burn_sev_sampling(burn_severity_14, gde_14) not enough data
burn_severity_78_sample <- burn_sev_sampling(burn_severity_78, gde_78)
burn_severity_80_sample <- burn_sev_sampling(burn_severity_80, gde_80)
burn_severity_81_sample <- burn_sev_sampling(burn_severity_81, gde_81)
burn_severity_85_sample <- burn_sev_sampling(burn_severity_85, gde_85)
```

## Bootstrapping
```{r}
eco_region_codes_burnsev <- c(1, 4, 5, 6, 78, 8, 80, 81, 85, 9) # remove 7, 13, and 14 due to insufficient data
eco_regions_names_burnsev <- eco_regions_names %>% 
  filter(us_l3code %in% eco_region_codes_burnsev)
eco_regions_names_burnsev_insuff <- eco_regions_names %>%
  filter(!us_l3code %in% eco_region_codes_burnsev)
eco_regions_geoms_burnsev <- eco_regions_geoms %>%
  filter(us_l3code %in% eco_region_codes_burnsev)

eco_regions_geoms_burnsev_insuff <- eco_regions_geoms %>%
  filter(!us_l3code %in% eco_region_codes_burnsev) %>%
  inner_join(eco_regions_names_burnsev_insuff) %>%
  mutate(sample = NA,
         confidence_interval_lower = NA, 
         confidence_interval_upper = NA, 
         p_value = NA, 
         result = "Insufficient Data"
         )
  
burn_severity_sample_list <- list()

for(i in 1:10){
  obj <- (paste0('burn_severity_', eco_region_codes_burnsev[i], '_sample'))
  burn_severity_sample_list <- c(burn_severity_sample_list, obj)
}

burn_severity_bootstrap <- bootstrap_results(burn_severity_sample_list) 

burn_severity_results <- burn_severity_bootstrap %>%
  bind_cols(eco_regions_names_burnsev) %>%
  inner_join(eco_regions_geoms_burnsev, by = "us_l3code") %>% 
  mutate(result = case_when(
    p_value < 0.05 & confidence_interval_upper < 0 ~ "Reject H0 - Incr. Burn Severity",
    p_value < 0.05 & confidence_interval_lower > 0 ~ "Reject H0 - Decr. Burn Severity",
    p_value > 0.05 ~ "Cannot Reject H0")) %>%
  bind_rows(eco_regions_geoms_burnsev_insuff)

burn_severity_results_sp <- burn_severity_results %>%
  subset(select = c("p_value","result", "geometry")) %>%
  st_as_sf(crs = crs_ca$wkt) 
```

## Map showing results from difference in mean test
```{r}
# grey is #EDF0F0
# red is #9D2A38
# teal is #2A9D8F
# cant reject is #C8D5D3
color_palette <- c("Reject H0 - Incr. Burn Severity" = "#9D2A38",
                   "Reject H0 - Decr. Burn Severity" = "#2A9D8F",
                   "Cannot Reject H0" = "#C8D5D3", 
                   "Insufficient Data" = "#EDF0F0")

tm_shape(burn_severity_results_sp) + 
  tm_polygons + 
  tm_fill(col = "result", palette = color_palette) + 
  tm_borders(col = "white", 
             lwd = 0.5) + 
  tm_layout(main.title = "",
            main.title.size = 0.8,
            frame = FALSE) + 
  tm_legend(position = c(1, -1))
```

## Histograms
```{r}
burn_severity_histogram <- function(sample_data, eco_region_code){
  
  # Create GDE dataframe for title
  df <- data.frame(
    code = c(0, 1, 4, 5, 6, 78, 8, 80, 81, 85, 9),
    region = c(
      "",
      "Coast Range",
      "Cascades",
      "Sierra Nevada",
      "Central California Foothills and Coastal Mountains",
      "Central California Valley",
      "Klamath Mountains/California High North Coast Range",
      "Southern California Mountains",
      "Northern Basin and Range",
      "Sonoran Basin and Range",
      "Southern California/Northern Baja Coast",
      "Eastern Cascades Slopes and Foothills"
    ),
    stringsAsFactors = FALSE
  )
  
  eco_region_name <- df$region[df$code == eco_region_code]
  
  plot_title = paste0(eco_region_name, ": Relative Frequency of Burn Severities in GDEs and Non-GDEs")
  
  fire_count_gdes <- sample_data %>%
    filter(gde == 1) %>%
    group_by(value) %>%
    summarize(proportion_1 = n() / nrow(.) * 100)
  
  fire_count_nongdes <- sample_data %>%
    filter(gde == 0) %>%
    group_by(value) %>%
    summarize(proportion_0 = n() / nrow(.) * 100)
  
  fire_count_histogram <- inner_join(fire_count_gdes, fire_count_nongdes, by = "value") %>%
    rename("GDE" = "proportion_1", 
           "NonGDE" = "proportion_0") %>% 
    pivot_longer(cols = c(2,3),
                 names_to = "gde_status",
                 values_to = "proportion"
    )
  
  ggplot(fire_count_histogram, aes(x = value, y = proportion, fill = gde_status)) + 
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_manual(values = c("#A3B18A", "#DDA15E")) +
    labs(x = "Fire Count", 
         y = "Relative Frequency (%)",
         fill = "GDE Status",
         title = plot_title) + 
    theme_classic() + 
    theme(legend.position = 'none',
          plot.title = element_text(hjust = 0.5,
                                    size = 15),
          axis.text = element_text(size = 13,
                                   color = 'black'),
          axis.title = element_text(size = 15,
                                    color = 'black'),
          axis.title.x = element_text(vjust = -1.1),
          axis.text.x = element_text(vjust = -1.5)) + 
    scale_y_continuous(expand = c(0,0)) + scale_x_continuous(expand = c(0,0))
}


```

