---
title: "Fire Count Raster"
author: "Jillian Allison"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r}
library(tidyverse)
library(sf)
library(here)
library(janitor)
library(raster)
library(rgdal)
library(fasterize)
library(exactextractr)
library(dplyr)
```

```{r}
# set working directory
setwd(here('~/Desktop/MEDS/Capstone/aquafire'))
```

## Load ecoregions and counties shapefiles 
```{r}
crs_ca <- st_crs(3310)

# Load in ecoregions shapefile
eco_regions <- read_sf(here::here('data', 'ca_eco_l3')) %>% 
  janitor::clean_names()  %>% 
  st_transform(crs_ca) %>% 
  rename(region = us_l3name)  %>%
  mutate(code = as.numeric(us_l3code))
eco_regions_names <- as.data.frame(eco_regions) %>%
  subset(select = c("us_l3code", "region"))

# Read in CA county boundaries
ca_counties <- read_sf(here::here('data', 'California_County_Boundaries', 'cnty19_1.shp')) %>% 
  janitor::clean_names() %>% 
  st_transform(crs_ca) # NAD27 / California Albers
```

## Load fire perimeters data and filter, set CRS, crop to California 
```{r}
fire_perimeters_all <- sf::read_sf(here('data', 
                                        'California_Fire_Perimeters_all')) %>% 
  clean_names() %>% 
  st_transform(crs_ca) %>% # NAD27 / California Albers
  mutate(val = 1) %>% # Create a column containing a value: this means each fire has a value of 1
  filter(year >= 1950) # Filter for only fires after 1950: this is when the dataset becomes reliable. 

# Crop to California
fires_ca <- st_crop(fire_perimeters_all, ca_counties) 
```

## Create an empty raster to use as a template
This valueless raster has a resolution of 30 x 30. We'll be using it to rasterize the fire perimeters layer. 
```{r}
## Set extent
ca_ext <- extent(ca_counties)
## Create valueless raster layer w/ 30 (unit is something) resolution 
ca_rast <- raster(ca_ext, resolution = 30)
crs(ca_rast) <- crs_ca$wkt
```

```{r}
# Use fasterize::fasterize() to convert polygon wildfires to raster layer based on raster template ## FUTURE US: REMEMBER TO CHECK FOR DUPLICATES!!!!! 
fire_count_raster <- fasterize(st_collection_extract(fires_ca, "POLYGON"), ca_rast, field = "val", fun = "sum", background = 0)
# Mask to California's boundary
fire_count_raster_masked <- mask(fire_count_raster, ca_counties) 
crs(fire_count_raster_masked) <- crs_ca$wkt

# Write raster file
#writeRaster(fire_count_raster_masked, "raster_output/firecount_updated430.tif", overwrite = TRUE)
```

```{r}
# Making a dataframe containing GDEs, which ecoregion they're in, and what the max fire count was in that GDE. Not quite what I was trying to make though... just had to see it to realize that. 
## I don't think this is really necessary, but I just wanna keep the code just in case. 
sb_county <- ca_counties %>% 
  filter(county_nam == "Santa Barbara")

gde_polygons_val <- st_make_valid(gde_polygons)

sb_gdes <- st_intersection(gde_polygons_val, sb_county)

sb_ecoregions <- st_intersection(eco_regions, sb_county)

sb_gdes_eco <- st_join(sb_gdes, sb_ecoregions) %>%
  subset(select = c(1:12, 31, 32, 42))

sb_fire_count_extract <- exactextractr::exact_extract(sb_fire_count, sb_gdes_eco, fun = "max")

sb_gde_eco_fc <- sb_gdes_eco %>%
  mutate(fire_count = sb_fire_count_extract)
```


```{r}
# The other one was weird, I'm gonna try with a new sb county shapefile 
sb_county <- read_sf(here::here('data', 'California_County_Boundaries', 'cnty19_1.shp')) %>% 
  janitor::clean_names() %>% 
  filter(objectid == 42) %>%
  st_transform(crs_ca) 

# Crop existing rasters to SB County Boundary
sb_fire_count <- crop(fire_count_raster_masked, sb_county)
sb_gde_raster <- crop(gde_masked, sb_county)
sb_eco_regions <- crop(eco_region_raster, sb_county) # ecoregion raster is in NLCD_Ecoregions_Layers.Rmd file 

# Make data frame containing how many times each cell has burned and a data frame that says if a cell is a GDE (1) or not (0) --- this does not contain info on ecoregion. BASIC REGRESSION INCOMING! 
sb_fire_count_df <- as.data.frame(sb_fire_count)

sb_gde_df <- as.data.frame(sb_gde_raster) 

sb_eco_regions_df <- as.data.frame(sb_eco_regions)

# Create index column in each DF for merging
cell_number <- 1:nrow(sb_fire_count_df)

sb_fire_count_df <- sb_fire_count_df %>%
  mutate(cell = cell_number)

sb_gde_df <- sb_gde_df %>%
  mutate(cell = cell_number,
         layer = as.factor(layer))

sb_eco_regions_df <- sb_eco_regions_df %>%
  mutate(cell = cell_number)

# Merge
sb_gde_fc_df <- merge(sb_fire_count_df, sb_gde_df, by = "cell") %>% 
  rename(fire_count = layer.x,
         gde = layer.y)

sb_gde_fc_er_df <- merge(sb_gde_fc_df, sb_eco_regions_df, by = "cell") %>%
  mutate(us_l3code = as.factor(layer)) %>%
  subset(select = -c(layer))%>%
  left_join(., eco_regions_names, by = "us_l3code")

# Group by ecoregion 

# Central CA Foothills and Coastal Mountains
gde_fc_6 <- sb_gde_fc_er_df %>%
  filter(us_l3code == 6)

# Central CA Valley
gde_fc_7 <- sb_gde_fc_er_df %>%
  filter(us_l3code == 7)

# Southern CA Mountains
gde_fc_8 <- sb_gde_fc_er_df %>%
  filter(us_l3code == 8)

# Southern CA / Northern Baja Coast
gde_fc_85 <- sb_gde_fc_er_df %>%
  filter(us_l3code == 85)


# Regress
summary(lm(fire_count ~ gde, data = sb_gde_fc_er_df))

## Central CA Foothills and Coastal Mountains
summary(lm(fire_count ~ gde, data = gde_fc_6)) 

## Central CA Valley 
## THERE WERE 0 FIRES HERE SO REMEMBER TO REMOVE AREAS THAT DON'T BURN! 
summary(lm(fire_count ~ gde, data = gde_fc_7)) 

## Southern CA Mountains
summary(lm(fire_count ~ gde, data = gde_fc_8)) 

## Southern CA / Northern Baja Coast
summary(lm(fire_count ~ gde, data = gde_fc_85))
```

Boxplot 
```{r}
ggplot(sb_gde_fc_er_df, aes(x = region, y = fire_count, fill = gde)) +
  geom_boxplot() +
  labs(x = "Ecoregion", y = "Fire count", fill = "GDE") +
  scale_fill_discrete(name = "GDE", labels = c("Non-GDE", "GDE")) +
  theme_bw()
```

## Dividing raster into ecoregions
I also wrote a function to do this based on this code, so hopefully this step is super easy! 

function is called raster_to_ecoregions
```{r}
# Crop raster to extent of shapefile
fc_cropped <- crop(fire_count_raster_masked, eco_regions)

# Convert shapefile to SpatialPolygonsDataFrame
eco_regions_sp <- as_Spatial(eco_regions) %>%
  SpatialPolygonsDataFrame(., data.frame(ID=1:nrow(eco_regions)))

# Loop through each eco region and crop raster to that region
for(i in 1:nrow(eco_regions)){
  # Subset shapefile to current eco region
  eco_region_subset <- eco_regions_sp[eco_regions_sp$ID == i,]
  
  # Crop raster to extent of eco region
  fc_eco_crop <- crop(fc_cropped, eco_region_subset)
  
  fc_subset <- mask(fc_eco_crop, eco_region_subset)
  
  # Write out raster to file
  writeRaster(fc_subset, filename = paste0("raster_output/fire_count_", eco_regions$us_l3code[i], ".tif"), format = "GTiff", overwrite = TRUE)
}

```

```{r}
fire_count_final <- remove_ag_urban(fire_count_raster_masked) # TAKES BETWEEN 15 - 30 MINUTES TO RUN! FYI 

writeRaster(fire_count_final, "raster_output/fire_count_final.tif")
```

```{r}
raster_to_ecoregions(fire_count_final, "fire_count", write_to_file = FALSE)
```

## Actual Stats
```{r}
# Coast Range 
fire_count_1_sample <- stratified_sample(fire_count_1, gde_1, n = 1000)
fire_count_1_gdes <- fire_count_1_sample %>%
  filter(gde == 1)
fire_count_1_nongdes <- fire_count_1_sample %>%
  filter(gde == 0)

fire_count_1_stats <- fire_count_stats(fire_count_1_sample)


# Plot density curves for GDEs and non-GDEs
library(ggplot2)
# Create violin plot
ggplot(fire_count_1_sample, aes(x = gde, y = value, fill = gde)) +
  geom_violin(trim = FALSE) +
  scale_fill_manual(values = c("springgreen4", "dodgerblue"), name = "GDE Status") + 
  xlab("GDE Status") +
  ylab("Fire Count") +
  ggtitle("Distribution of Fire Counts in Non-GDEs vs GDEs") + 
  theme_minimal()


ggplot(fire_count_1_sample, aes(x = value, color = gde)) + 
  stat_ecdf() +
  scale_color_manual(values = c("springgreen4", "dodgerblue"), name = "GDE Status") +
  xlab("Fire Count") +
  ylab("Cumulative Proportion") +
  ggtitle("Cumulative Distribution of Fire Counts in GDEs vs Non-GDEs") + 
  theme_minimal()

```



