---
title: "Jillian Capstone"
author: "Jillian Allison"
date: "2023-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r}
library(tidyverse)
library(sf)
library(here)
library(janitor)
```

```{r}
# set working directory
setwd(here('~/Desktop/MEDS/Capstone/aquafire'))
```

## Load Boundaries
```{r}
# Load in ecoregions shapefile
eco_regions <- read_sf(here::here('data', 'ca_eco_l3')) %>% 
  janitor::clean_names()  %>% 
  st_transform("EPSG: 3309") %>% 
  rename(region = us_l3name)

# Read in CA county boundaries
ca_counties <- read_sf(here::here('data', 'CA_Counties')) %>% 
  janitor::clean_names() %>% 
  st_transform("EPSG: 3309") # NAD27 / California Albers

# Filter sb county
sb_county <- ca_counties %>% 
  filter(name == "Santa Barbara")
```

## Load GDE Data
```{r}
# veg_layer <- st_read(dsn = here('data', 'i02_NaturalCommunitiesCommonlyAssociatedwithGroundwater_v2_0.gdb'),
#                      layer = 'i02_NCCAG_Vegetation_2_0') %>% 
#   st_transform("EPSG: 3309")
# 
# wetland_layer <- st_read(dsn = here('data', 'i02_NaturalCommunitiesCommonlyAssociatedwithGroundwater_v2_0.gdb'),
#                      layer = 'i02_NCCAG_Wetlands_2_0')
# 



# # read in groundwater dependent ecosystem shapefile + clean names
# gde_veg <- read_sf(here('data',
#                         'i02_naturalcommunitiescommonlyassociatedwithgroundwater',
#                         'i02_NCCAG_Vegetation.shp')) %>% 
#   clean_names() %>% 
#   st_transform("EPSG: 3309") # NAD27 / California Albers
# 
# gde_wetlands <- read_sf(here('data','i02_naturalcommunitiescommonlyassociatedwithgroundwater','i02_NCCAG_Wetlands.shp')) %>% 
#   clean_names()  %>% 
#   st_transform("EPSG: 3309") # NAD27 / California Albers
# 
# gde <- st_join(gde_veg, gde_wetlands)
# 
# gde_ndvi <- st_read(here('data', 'i02_NaturalCommunitiesCommonlyAssociatedwithGroundwater_v2_0.gdb')) %>% 
#   st_transform("EPSG: 3309") %>% 
#   clean_names() %>% 
#   st_cast(gde_ndvi, "MULTIPOLYGON")
```

## Load Fire Data
```{r}
fire_perimeters_all <- sf::read_sf(here('data', 
                                        'California_Fire_Perimeters_all')) %>% 
  clean_names() %>% 
  st_transform("EPSG: 3309") # NAD27 / California Albers
```

```{r}
fires_sb <- st_crop(fire_perimeters_all, sb_county) %>%
  mutate(val = 1)
```

```{r}
library(raster)
ext <- extent(fires_sb)

r <- raster(ext, resolution = 30) 
crs(r) <- "+init=epsg:3309"
```

```{r}
library(fasterize)

# irrelevant, was just trying something 
hollister <- fires_sb %>%
  filter(fire_name == "HOLLISTER")

hollister_r <- fasterize(hollister, r)
```

```{r}
# CHECK FOR DUPLICATES!!!!! 
testr <- fasterize(st_collection_extract(fires_sb, "POLYGON"), r, field = "val", fun = "sum", background = 0)
```

```{r}
plot(testr)
```

## FIRE COUNTS FOR ALL OF CALIFORNIA 
```{r}
# Load in fire perimeter data (did earlier), and mutate to create a value column. each fire has a value of 1. This will be used for summing fires. 
fire_perimeters_all <- fire_perimeters_all %>%
  mutate(val = 1) %>%
  filter(year >= 1950)

# Crop to California
fires_ca <- st_crop(fire_perimeters_all, ca_counties) %>%
  mutate(val = 1)

# Create raster layer template for CA 
library(raster)
## Set extent
ca_ext <- extent(fires_ca)
## Create valueless raster layer w/ 30 (unit is something) resolution 
ca_rast <- raster(ca_ext, resolution = 30) # check that this is 30 *meters* 
crs(ca_rast) <- "+init=epsg:3309" # set crs

# Use fasterize::fasterize() to convert polygon wildfires to raster layer based on raster template ## FUTURE US: REMEMBER TO CHECK FOR DUPLICATES!!!!! 
fire_count_raster <- fasterize(st_collection_extract(fires_ca, "POLYGON"), ca_rast, field = "val", fun = "sum", background = 0)

plot(fire_count_raster)
```


## TSLF Based on YEAR column for all of California 
Data was most recently updated in November 2022. 
Steps: 
1) Create a raster layer with each cell = most recent fire's year in that cell. if no fire, then value = NA
2) Create RL of 2022 - recent fire 

```{r}
fires_ca <- fires_ca %>%
  mutate(year = as.numeric(year))

# Use fasterize::fasterize() to create a raster layer where each cell is equal to the maximum year value for that cell. 
most_recent_raster <- fasterize(st_collection_extract(fires_ca, "POLYGON"), ca_rast, field = "year", fun = "max", background = NA)
```

```{r}
plot(most_recent_raster)
```

```{r}
# Subtract the raster layer containing the most recent fire from 2022- this will give the number of years since the most recent fire in each cell. 
tslf_raster <- 2022 - most_recent_raster
```


### For sharing plots
```{r}
plot(fire_count_raster)
```
pick cutoff date (1950) vvv there too

```{r}
plot(tslf_raster)
```
* examine seasonal differences (try grouping by season) 

```{r}
# Lets mask the layers we made here
fire_count_raster_masked <- mask(fire_count_raster, ca_counties)

tslf_raster_masked <- mask(tslf_raster, ca_counties)
```

```{r}
# Write files 
# writeRaster(fire_count_raster_masked, "firecount.tif")
# writeRaster(tslf_raster_masked, "tslf.tif")
```


### GDE Layer 
```{r}
# Read in GDE data
gde_data <- sf::read_sf("data/gde") 

gde_data <- gde_data %>% 
  mutate(gde = 1)
```

```{r}
gde_raster <- fasterize::fasterize(st_collection_extract(gde_data, "POLYGON"), ca_rast, field = "gde", fun = "max", background = NA)

gde_masked <- mask(gde_raster, ca_counties)
```


```{r}
writeRaster(gde_masked, "gde_boundaries.tif")
```




