---
title: "Raster Layer Code"
author: "Jillian Allison"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries and Functions

```{r}
library(tidyverse)
library(sf)
library(here)
library(janitor)
library(raster)
library(rgdal)
library(fasterize)
library(exactextractr)
library(dplyr)
library(terra)
library(tmap)
source("Functions/remove_ag_urban.R")
source("Functions/raster_to_ecoregions.R")
source("Functions/stratified_sample.R")
```

# Read in data

## Load ecoregions and counties shapefiles

```{r}
crs_ca <- st_crs(3310)

# Load in ecoregions shapefile
eco_regions <- read_sf(here::here('data', 'ca_eco_l3')) %>% 
  janitor::clean_names()  %>% 
  st_transform(crs_ca) %>% 
  rename(region = us_l3name)  %>%
  mutate(code = as.numeric(us_l3code))
# Create file containing eco region names 
eco_regions_names <- as.data.frame(eco_regions) %>%
  subset(select = c("us_l3code", "region"))

# Read in CA county boundaries
ca_counties <- read_sf(here::here('data', 'California_County_Boundaries', 'cnty19_1.shp')) %>% 
  janitor::clean_names() %>% 
  st_transform(crs_ca) # NAD27 / California Albers
```

## Load fire perimeters data and filter, set CRS, crop to California

```{r}
fire_perimeters_all <- sf::read_sf(here('data', 
                                        'California_Fire_Perimeters_all')) %>% 
  clean_names() %>% 
  st_transform(crs_ca) %>% # NAD27 / California Albers
  mutate(val = 1) %>% # Create a column containing a value: this means each fire has a value of 1
  filter(year >= 1950) # Filter for only fires after 1950: this is when the dataset becomes reliable. 

# Crop to California
fires_ca <- st_crop(fire_perimeters_all, ca_counties) 
```

## Load GDE Data

```{r}
# Read in GDE data
gde_data <- sf::read_sf(here::here("data", "gde") )

gde_polygons_raster <- gde_data %>%
  st_transform(crs_ca) %>%
  filter(!grepl("Permanently Flooded", WETLAND_NA)) %>% # Removes permanently submerged GDEs
  mutate(gde = 1)

gde_polygons_shiny <- gde_data %>%
  st_transform(crs_ca) %>%
  filter(!grepl("Permanently Flooded", WETLAND_NA)) %>% # Removes permanently submerged GDEs
  mutate(area_m2 = as.numeric(st_area(geometry))) %>%
  filter(area_m2 > 8093) %>% # sq meters, 2 acres 
  st_simplify(preserveTopology = TRUE, dTolerance = 5) # Simplify remaining polygons to reduce load time in Shiny App
```

## Create an empty raster to use as a template

This valueless raster has a resolution of 30 x 30. We'll be using it to rasterize the fire perimeters layer.

```{r}
crs_ca <- st_crs(3310)
## Set extent
ca_ext <- extent(ca_counties)
## RASTER OBJECT Create valueless raster layer w/ 30 (unit is something) resolution 
ca_rast <- raster(ca_ext, resolution = 30)
crs(ca_rast) <- crs_ca$wkt

## TERRA OBJECT Create valueless raster layer w/ 30 (unit is something) resolution 
ca_rasterra <- raster(ca_ext, resolution = 30) %>%
  terra::rast()
crs(ca_rasterra) <- crs_ca$wkt 
```

## Groundwater Dependent Ecosystems

### GDE Raster

```{r}
# Fasterize to rasterize, extent to match others 
gde_raster <- fasterize::fasterize(st_collection_extract(gde_polygons_raster, "POLYGON"), ca_rast, field = "gde", fun = "max", background = 0)

# Mask to california
gde_masked <- mask(gde_raster, ca_counties)

# Remove agricultural and urban areas
gde_final <- remove_ag_urban(input_raster_layer = gde_masked, nlcd_raster_layer = nlcd_rasterlayer)

# Create raster file 
writeRaster(gde_final, here::here("raster_output", "gde_final.tif"))

# Create GDE Objects divided into ecoregions 
raster_to_ecoregions(raster_layer = gde_final, file_name = "gde", write_to_file = TRUE)
```

### GDE Polygons with Summary Data

```{r}
# Remove polygons that are ag or urban from the GDE polygon data
# First, extract land use data 
gdepoly_nlcd <- exactextractr::exact_extract(nlcd_rasterlayer, gde_polygons_shiny, fun = "majority")

# Join and filter
nlcd_vals_to_remove <- c(11, 21, 22, 23, 24, 81, 82)

gde_polygons_shiny <- gde_polygons_shiny %>%
  mutate(nlcd = gdepoly_nlcd) %>%
  filter(!nlcd %in% nlcd_vals_to_remove)

st_write(gde_polygons_shiny, "raster_output/gde_polygons_rm_ag_urb.shp")
```

```{r}
# Find the maximum number of times that a GDE polygon has burned since 1950 using the fire count raster layer. Find maximum number of times a cell has burned in that GDE
gdepoly_fire_count <- exactextractr::exact_extract(fire_count_final, gde_polygons_shiny, fun = "max")

# Find most recent TSLF in each GDE polygon
gdepoly_tslf <- exactextractr::exact_extract(tslf_final, gde_polygons_shiny, fun = "min")


# Find average modeled fire threat in each GDE polygon
gdepoly_fire_threat <- exactextractr::exact_extract(fire_threat_final, gde_polygons_shiny, fun = "mean")


# Find mode burn severity in GDE polygon 
gdepoly_burn_severity <- exactextractr::exact_extract(burn_severity_final, gde_polygons_shiny, fun = "mean")



gdepoly_perc_burned <- exactextractr::exact_extract(perc_burned_final, gde_polygons_shiny, fun = "mean")
```

```{r}
# Join data! 
gde_summary_shiny <- gde_polygons_shiny %>%
  mutate(max_fire_count = gdepoly_fire_count, 
         min_tslf = gdepoly_tslf, 
         avg_fire_threat = gdepoly_fire_threat, 
         avg_fire_severity = gdepoly_burn_severity) %>%
  mutate(area_km2 = (area_m2 / (1E6)),
         area_km2 = round(area_km2, digits = 3),
         max_fire_count = round(max_fire_count, digits = 2), 
         min_tslf = round(min_tslf, digits = 2), 
         avg_fire_threat = round(avg_fire_threat, digits = 2), 
         avg_fire_severity = round(avg_fire_severity, digits = 2), 
         avg_fire_threat = na_if(avg_fire_threat, NaN),
         avg_fire_severity = na_if(avg_fire_severity, NaN))

# adding fire_severity
gde_summary_shiny <- gde_summary_shiny %>%
  mutate(avg_fire_severity = gdepoly_burn_severity) %>%
  mutate(avg_fire_severity = na_if(avg_fire_severity, NaN)) %>%
  subset(select = -c(avg_fr_s, avg_fr_sv)) %>%
  st_as_sf(crs = crs_ca$wkt)

st_write(gde_summary_shiny, "raster_output/gde_summary_shiny.shp", append = F)
  
```

```{r}
# Perform spatial join to assign eco region IDs to GDE polygons
gde_summ_eco <- st_join(gde_summary_shiny, eco_regions) 

gde_summ_eco_df <- as.data.frame(gde_summ_eco)

ecoregion_codes <- as.numeric(unique(gde_summ_eco$us_l3code))

for(id in ecoregion_codes){
  gdes_in_ecoregion <- gde_summ_eco %>%
    filter(us_l3code == id) 
  
  output_file <- paste0("raster_output/gde_shiny_summary_ecoregions/gde_summary_", id, ".shp")
  
  st_write(gdes_in_ecoregion, output_file)
}

st_write(gde_summ_eco, "raster_output/gde_shiny_summary_ecoregions/gde_summary_ALL.shp")
```

### GDE Polygons by Eco Region

```{r}
# Perform spatial join to assign eco region IDs to GDE polygons
gde_eco <- st_join(gde_polygons_shiny, eco_regions) 

ecoregion_codes <- as.numeric(unique(gde_eco$us_l3code))

for(id in ecoregion_codes){
  gdes_in_ecoregion <- gde_eco %>%
    filter(us_l3code == id) %>%
    select(POLYGON_ID, WETLAND_NA, VEGETATION, DOMINANT_S, DOMINANT_C, area, us_l3code, region, geometry)
  output_file <- paste0("raster_output/gde_eco_regions_simplified/gde_polygon_", id, ".shp")
  
  st_write(gdes_in_ecoregion, output_file)
}
```

# Eco Region Raster

```{r}
# Fasterize to rasterize, extent to match others 
eco_region_raster <- fasterize(eco_regions, ca_rast, field = "us_l3code")
```

# NLCD

```{r}
# I'm using the fed data package because NLCD data was in .img format and .... ew 
library(FedData)

nlcd <- FedData::get_nlcd(template = ca_rasterra, year = 2019, label = "ca", dataset = "landcover", landmass = "L48", extraction.dir = "data") %>%
  rast() %>%
  terra::project(y = crs_ca$wkt, method = "near") # takes several minutes to run, but it will! 


nlcd_resampled <- terra::resample(nlcd, ca_rasterra, method = "near") # uses nearest neighbors approach for for categorical values 

nlcd_masked <- mask(nlcd_resampled, ca_counties)

nlcd_rasterlayer <- raster(nlcd_masked)

# raster::writeRaster(nlcd_rasterlayer, "raster_output/nlcd.tif", overwrite = TRUE)

# nlcd_colors() and pal_nlcd() return legend and color palette
# nlcd_colors()
# pal_nlcd()
```

# Fire Count

```{r}
# Use fasterize::fasterize() to convert polygon wildfires to raster layer based on raster template ## FUTURE US: REMEMBER TO CHECK FOR DUPLICATES!!!!! 
fire_count_raster <- fasterize(st_collection_extract(fires_ca, "POLYGON"), ca_rast, field = "val", fun = "sum", background = 0)
# Mask to California's boundary
fire_count_raster_masked <- mask(fire_count_raster, ca_counties) 
crs(fire_count_raster_masked) <- crs_ca$wkt
```

## Dividing raster into ecoregions

```{r}
fire_count_final <- remove_ag_urban(fire_count_raster_masked) # TAKES BETWEEN 15 - 30 MINUTES TO RUN! FYI 

# raster_to_ecoregions(fire_count_final, "fire_count", write_to_file = TRUE)

# writeRaster(fire_count_final, "raster_output/fire_count_final.tif")
```

# TSLF

```{r}
# Use fasterize::fasterize() to create a raster layer where each cell is equal to the maximum year value for that cell. 
most_recent_raster <- fasterize(st_collection_extract(fires_ca, "POLYGON"), ca_rast, field = "year", fun = "max", background = NA)

# Subtract the raster layer containing the most recent fire from 2022- this will give the number of years since the most recent fire in each cell. 
tslf_raster <- 2022 - most_recent_raster

# Mask to California's boundary
tslf_raster_masked <- mask(tslf_raster, ca_counties)
crs(tslf_raster_masked) <- crs_ca$wkt

# Remove ag and urban areas 
# tslf_final <- remove_ag_urban(tslf_raster_masked) # TAKES BETWEEN 15 - 30 MINUTES TO RUN! FYI

raster_to_ecoregions(tslf_final, "tslf", write_to_file = TRUE)

# writeRaster(tslf_final, "raster_output/tslf_final.tif")
```

# Fire Threat

```{r}
fire_threat <- raster(here::here("data", "fire_threat.tif")) %>%
  terra::rast() %>%
  terra::project(y = crs_ca$wkt, method = "near")
  

fire_threat_resampled <- terra::resample(fire_threat, ca_rasterra, method = "near") # uses nearest neighbors approach for for categorical values 

fire_threat_masked <- mask(fire_threat_resampled, ca_counties)

fire_threat_rasterlayer <- raster(fire_threat_masked)

# Remove ag and urban areas 
fire_threat_final <- remove_ag_urban(fire_threat_rasterlayer) # TAKES BETWEEN 15 - 30 MINUTES TO RUN! FYI 

writeRaster(fire_threat_final, "raster_output/fire_threat_final.tif", overwrite = TRUE)

raster_to_ecoregions(fire_threat_final, "fire_threat", write_to_file = TRUE)
```

# Burn Severity
```{r}
burn_severity_final <- raster("raster_output/burn_severity_mode.tif")
burn_severity_final <- burn_severity_final %>%
  terra::rast() %>%
  terra::project(y = crs_ca$wkt, method = "near") %>%
  terra::resample(ca_rasterra, method = "near") %>%
  terra::mask(ca_counties) %>%
  raster()

burn_severity_final <- remove_ag_urban(burn_severity_final)
writeRaster(burn_severity_final, "raster_output/burn_severity_final.tif", overwrite = TRUE)
raster_to_ecoregions(burn_severity_final, "burn_severity", write_to_file = TRUE)
beep(sound = 3)
```

