---
title: "ecoregion_info"
author: "Meagan Brown"
date: "2023-05-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load necessary data
```{r}
gde <- read_sf(here::here('data', 'gde')) %>% 
  st_transform("EPSG: 3310")

# ecoregions data
eco_regions <- read_sf(here::here('data', 'ca_eco_l3')) %>% 
  janitor::clean_names()  %>% 
  st_transform("EPSG: 3310") %>% 
  rename(region = us_l3name)
```

# 1. COAST RANGE
```{r}
# Filter for just the Southern California/Northern Baja Coast Ecoregion
coast_range <- eco_regions %>% 
  filter(region == 'Coast Range')

# Enable interactive mode and switch to leaflet mode
tmap_mode("plot")

# Define the text information for the popup
popup_text <- "1. COAST RANGE"

# Add a column for the popup text to your data
coast_range$popup_text <- popup_text

# Plot the polygons with interactive click events and custom popup
tm_shape(coast_range) +
  tm_polygons(col = "#0081A7", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") +
   tm_layout(frame = FALSE)
```

# 4. CASCADES
```{r}
# Filter for just the Southern California/Northern Baja Coast Ecoregion
cascades <- eco_regions %>% 
  filter(region == 'Cascades')

# Enable interactive mode and switch to leaflet mode
tmap_mode("plot")

# Define the text information for the popup
popup_text <- "4. CASCADES"

# Add a column for the popup text to your data
cascades$popup_text <- popup_text

# Plot the polygons with interactive click events and custom popup
  tm_shape(cascades) +
  tm_polygons(col = "#8CB369", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") + tm_layout(frame = FALSE)
```

# 5. SIERRA NEVADA
```{r}
# Filter for just the Southern California/Northern Baja Coast Ecoregion
sierra_nevada <- eco_regions %>% 
  filter(region == 'Sierra Nevada')

# Enable interactive mode and switch to leaflet mode
tmap_mode("plot")

# Define the text information for the popup
popup_text <- "5. SIERRA NEVADA"

# Add a column for the popup text to your data
sierra_nevada$popup_text <- popup_text

# Plot the polygons with interactive click events and custom popup
  tm_shape(sierra_nevada) +
  tm_polygons(col = "#C0CB77", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") + tm_layout(frame = FALSE)
```

# 6. CENTRAL CALIFORNIA FOOTHILLS AND COASTAL MOUNTAINS
```{r}
# Filter for just the Southern California/Northern Baja Coast Ecoregion
central_cal_coast_mtn <- eco_regions %>% 
  filter(region == 'Central California Foothills and Coastal Mountains')

# Enable interactive mode and switch to leaflet mode
tmap_mode("plot")

# Define the text information for the popup
popup_text <- "6. CENTRAL CALIFORNIA FOOTHILLS AND COASTAL MOUNTAINS"

# Add a column for the popup text to your data
central_cal_coast_mtn$popup_text <- popup_text

# Plot the polygons with interactive click events and custom popup
  tm_shape(central_cal_coast_mtn) +
  tm_polygons(col = "#7FD6CB", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") + tm_layout(frame = FALSE)
```

# 7. CENTRAL CALIFORNIA VALLEY
```{r}
# Filter for just the Southern California/Northern Baja Coast Ecoregion
central_valley <- eco_regions %>% 
  filter(region == 'Central California Valley')

# Enable interactive mode and switch to leaflet mode
tmap_mode("plot")

# Define the text information for the popup
popup_text <- "7. CENTRAL CALIFORNIA VALLEY"

# Add a column for the popup text to your data
central_valley$popup_text <- popup_text

# Plot the polygons with interactive click events and custom popup
  tm_shape(central_valley) +
  tm_polygons(col = "lightgrey", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.5, 
              border.col = 'white', lwd = 0.3,
              lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right")  + tm_layout(frame = FALSE)
```

# 8. SOUTHERN CALIFORNIA MOUNTAINS
```{r}
# Filter for just the Southern California/Northern Baja Coast Ecoregion
socal_mtns <- eco_regions %>% 
  filter(region == 'Southern California Mountains')

# Enable interactive mode and switch to leaflet mode
tmap_mode("plot")

# Define the text information for the popup
popup_text <- "8. SOUTHERN CALIFORNIA MOUNTAINS"

# Add a column for the popup text to your data
socal_mtns$popup_text <- popup_text

# Plot the polygons with interactive click events and custom popup
  tm_shape(socal_mtns) +
  tm_polygons(col = "#F4E285", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") + tm_layout(frame = FALSE)
```

# 9. EASTERN CASCADE SLOPES AND FOOTHILLS
```{r}
# Filter for just the Southern California/Northern Baja Coast Ecoregion
east_cascade_foothills <- eco_regions %>% 
  filter(region == 'Eastern Cascades Slopes and Foothills')

# Enable interactive mode and switch to leaflet mode
tmap_mode("plot")

# Define the text information for the popup
popup_text <- "9. EASTERN CASCADE SLOPES AND FOOTHILLS"

# Add a column for the popup text to your data
east_cascade_foothills$popup_text <- popup_text

# Plot the polygons with interactive click events and custom popup
  tm_shape(east_cascade_foothills) +
  tm_polygons(col = "#F4C26F", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") + tm_layout(frame = FALSE)
```

# 13. CENTRAL BASIN AND RANGE
```{r}
# Filter for just the Southern California/Northern Baja Coast Ecoregion
central_basin_range <- eco_regions %>% 
  filter(us_l3code == '13')

# Enable interactive mode and switch to leaflet mode
tmap_mode("plot")

# Define the text information for the popup
popup_text <- "13. CENTRAL BASIN AND RANGE"

# Add a column for the popup text to your data
central_basin_range$popup_text <- popup_text

# Plot the polygons with interactive click events and custom popup
 # central basin and range
  tm_shape(central_basin_range) +
  tm_polygons(col = "#F4A259", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") + tm_layout(frame = FALSE)
```

# 14. MOJAVE BASIN AND RANGE
```{r}
# Filter for just the Southern California/Northern Baja Coast Ecoregion
mojave_basin_range <- eco_regions %>% 
  filter(us_l3code == '14')

# Enable interactive mode and switch to leaflet mode
tmap_mode("plot")

# Define the text information for the popup
popup_text <- "14. MOJAVE BASIN AND RANGE"

# Add a column for the popup text to your data
mojave_basin_range$popup_text <- popup_text

# Plot the polygons with interactive click events and custom popup
  tm_shape(mojave_basin_range) +
  tm_polygons(col = "#8CB369", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") + tm_layout(frame = FALSE)
```

# 78. KLAMATH MOUNTAINS AND CALIFORNIA HIGH NORTH COAST RANGE
```{r}
# Filter for just the Southern California/Northern Baja Coast Ecoregion
klamath_mtns <- eco_regions %>% 
  filter(us_l3code == '78')

# Enable interactive mode and switch to leaflet mode
tmap_mode("plot")

# Define the text information for the popup
popup_text <- "78. KLAMATH MOUNTAINS AND CALIFORNIA HIGH NORTH COAST RANGE"

# Add a column for the popup text to your data
klamath_mtns$popup_text <- popup_text

# Plot the polygons with interactive click events and custom popup
  tm_shape(klamath_mtns) +
  tm_polygons(col = "#A8986B", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") + tm_layout(frame = FALSE)
```

# 80. NORTHERN BASIN AND RANGE
```{r}
# Filter for just the Southern California/Northern Baja Coast Ecoregion
northern_basin_range <- eco_regions %>% 
  filter(us_l3code == '80')

# Enable interactive mode and switch to leaflet mode
tmap_mode("plot")

# Define the text information for the popup
popup_text <- "80. NORTHERN BASIN AND RANGE"

# Add a column for the popup text to your data
northern_basin_range$popup_text <- popup_text

# Plot the polygons with interactive click events and custom popup
tm_shape(northern_basin_range) +
  tm_polygons(col = "#829374", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") + tm_layout(frame = FALSE)
```

# 81. SONORAN BASIN AND RANGE
```{r}
# Filter for just the Southern California/Northern Baja Coast Ecoregion
sonoran_basin_range <- eco_regions %>% 
  filter(us_l3code == '81')

# Enable interactive mode and switch to leaflet mode
tmap_mode("plot")

# Define the text information for the popup
popup_text <- "81. SONORAN BASIN AND RANGE"

# Add a column for the popup text to your data
sonoran_basin_range$popup_text <- popup_text

# Plot the polygons with interactive click events and custom popup
  tm_shape(sonoran_basin_range) +
  tm_polygons(col = "#5B8E7D", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") + tm_layout(frame = FALSE)
```

# 85. SOUTHERN CALIFORNIA/NORTHERN BAJA COAST
```{r}
# Filter for just the Southern California/Northern Baja Coast Ecoregion
socal_baja_coast <- eco_regions %>% 
  filter(region == 'Southern California/Northern Baja Coast')

# Enable interactive mode and switch to leaflet mode
tmap_mode("plot")

# Define the text information for the popup
popup_text <- "85. SOUTHERN CALIFORNIA/NORTHERN BAJA COAST"

# Add a column for the popup text to your data
socal_baja_coast$popup_text <- popup_text

# Plot the polygons with interactive click events and custom popup
  tm_shape(socal_baja_coast) +
  tm_polygons(col = "#BC4B51", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") + tm_layout(frame = FALSE)
```

# ALL REGIONS TOGETHER

```{r}
tmap_mode("view")

# coast range
tm_shape(coast_range) +
  tm_polygons(col = "#0081A7", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") +
  
  # cascades
  tm_shape(cascades) +
  tm_polygons(col = "#8CB369", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") +
  
  # sierra nevada
  tm_shape(sierra_nevada) +
  tm_polygons(col = "#C0CB77", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") +
  
  # central cal mtns
  tm_shape(central_cal_coast_mtn) +
  tm_polygons(col = "#7FD6CB", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") +
  
  # socal mountains
  tm_shape(socal_mtns) +
  tm_polygons(col = "#F4E285", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") +
  
  # east cascade foothills
  tm_shape(east_cascade_foothills) +
  tm_polygons(col = "#F4C26F", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") +
  
  # central basin and range
  tm_shape(central_basin_range) +
  tm_polygons(col = "#F4A259", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") +
  
  # mojave basin and range
  tm_shape(mojave_basin_range) +
  tm_polygons(col = "#8CB369", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") +
  
  # klamath mtns
  tm_shape(klamath_mtns) +
  tm_polygons(col = "#A8986B", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") +
  
  # northern basin
  tm_shape(northern_basin_range) +
  tm_polygons(col = "#829374", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3, lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") +
  
  # sonoran basin
  tm_shape(sonoran_basin_range) +
  tm_polygons(col = "#5B8E7D", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") +
  
  # socal baja coast
  tm_shape(socal_baja_coast) +
  tm_polygons(col = "#BC4B51", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.55, 
              border.col = 'white', lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") +
  
  # central valley
  tm_shape(central_valley) +
  tm_polygons(col = "lightgrey", 
              title = "Region", 
              id = "popup_text", 
              alpha = 0.5, 
              border.col = 'white', lwd = 0.3,
              lwd = 0.3) +
  tm_layout(legend.outside = TRUE, 
            legend.outside.position = "right") 
```


