---
title: "Fire Count Analysis"
author: "Jillian Allison, Meagan Brown, Andre Dextre, Wade Sedgwick (MEDS 2023)"
date: "`r Sys.Date()`"
output: html_document
---

### Eco Region Key: 

   **code**                                           **region**
          1                                         Coast Range
         13                             Central Basin and Range
         14                              Mojave Basin and Range
          4                                            Cascades
          5                                       Sierra Nevada
          6  Central California Foothills and Coastal Mountains
          7                           Central California Valley
         78 Klamath Mountains/California High North Coast Range
          8                       Southern California Mountains
         80                            Northern Basin and Range
         81                             Sonoran Basin and Range
         85             Southern California/Northern Baja Coast
          9               Eastern Cascades Slopes and Foothills 
          

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages and Functions 
```{r}
library(tidyverse)
library(sf)
library(boot)
library(boot.pval)
library(here)
library(janitor)
library(raster)
library(tmap)
library(rgdal)
library(fasterize)
library(exactextractr)
library(dplyr)
library(gridExtra)
library(kableExtra)
source("Functions/remove_ag_urban.R")
source("Functions/raster_to_ecoregions.R")
source("Functions/stratified_sample.R")
source("Functions/fire_count_violin_plot.R")
source("Functions/fire_count_stats.R")
source("Functions/diff_in_means_bootstrap.R")
source("Functions/bootstrap_results.R")
source("Functions/fire_count_histogram.R")
source("Functions/burn_sev_sampling.R")
source("Functions/mann_whitney_result.R")
source("Functions/mann_whitney_test.R")
source("Functions/burn_severity_histogram.R")
```

### Necessary Objects 
To run this code, you will need to run the GDE and Fire Count sections of the "Spatial_Layers.rmd" file in the Spatial_Layers folder in this repository. That code requires the proper datasets- see the project repository on Dangermond's dataone.org repository to access that data. Once you have run that code, you will be able to proceed with this analysis. 
If you are using new data, you may read in .tif files for GDEs and fire count data if they share the same resolution, CRS, and extent. Use compareRaster() to check if your raster input layers align. 
```{r}
# For eco region names / codes 
crs_ca <- st_crs(3310)
# Load in ecoregions shapefile
eco_regions <- read_sf(here::here('data', 'ca_eco_l3')) %>% 
  janitor::clean_names()  %>% 
  st_transform(crs_ca) %>% 
  rename(region = us_l3name)  %>%
  mutate(code = as.numeric(us_l3code))
# Create file containing eco region names 
eco_regions_names <- as.data.frame(eco_regions) %>%
  subset(select = c("us_l3code", "region"))
eco_regions_geoms <- as.data.frame(eco_regions) %>%
  subset(select = c("us_l3code", "geometry"))
eco_region_codes <- c(1, 13, 14, 4, 5, 6, 7, 78, 8, 80, 81, 85, 9)

gde_1 <- raster("raster_output/gde_1.tif")
gde_4 <- raster("raster_output/gde_4.tif")
gde_5 <- raster("raster_output/gde_5.tif")
gde_6 <- raster("raster_output/gde_6.tif")
gde_7 <- raster("raster_output/gde_7.tif")
gde_8 <- raster("raster_output/gde_8.tif")
gde_9 <- raster("raster_output/gde_9.tif")
gde_13 <- raster("raster_output/gde_13.tif")
gde_14 <- raster("raster_output/gde_14.tif")
gde_78 <- raster("raster_output/gde_78.tif")
gde_80 <- raster("raster_output/gde_80.tif")
gde_81 <- raster("raster_output/gde_81.tif")
gde_85 <- raster("raster_output/gde_85.tif")

fire_count_1 <- raster("raster_output/fire_count_1.tif")
fire_count_4 <- raster("raster_output/fire_count_4.tif")
fire_count_5 <- raster("raster_output/fire_count_5.tif")
fire_count_6 <- raster("raster_output/fire_count_6.tif")
fire_count_7 <- raster("raster_output/fire_count_7.tif")
fire_count_8 <- raster("raster_output/fire_count_8.tif")
fire_count_9 <- raster("raster_output/fire_count_9.tif")
fire_count_13 <- raster("raster_output/fire_count_13.tif")
fire_count_14 <- raster("raster_output/fire_count_14.tif")
fire_count_78 <- raster("raster_output/fire_count_78.tif")
fire_count_80 <- raster("raster_output/fire_count_80.tif")
fire_count_81 <- raster("raster_output/fire_count_81.tif")
fire_count_85 <- raster("raster_output/fire_count_85.tif")
```

## Obtain Sample Data
This code uses the stratified_sample() function in the "Functions" folder of this repository. The function inputs are the relevant fire layer, the corresponding GDE layer, and the number of points you would like to sample. 
In the code below, we take a stratified random sample of 1000 points within and outside of GDEs in each ecoregion. Each point has a buffer distance of 1000 m / 1 km, meaning that points will be at least 1 km apart so the user can assume independence. 

```{r}
set.seed(123) 
fire_count_1_sample <- stratified_sample(fire_count_1, gde_1, n = 1000)
fire_count_4_sample <- stratified_sample(fire_count_4, gde_4, n = 1000)
fire_count_5_sample <- stratified_sample(fire_count_5, gde_5, n = 1000)
fire_count_6_sample <- stratified_sample(fire_count_6, gde_6, n = 1000)
fire_count_7_sample <- stratified_sample(fire_count_7, gde_7, n = 1000)
fire_count_8_sample <- stratified_sample(fire_count_8, gde_8, n = 1000)
fire_count_9_sample <- stratified_sample(fire_count_9, gde_9, n = 1000)
fire_count_13_sample <- stratified_sample(fire_count_13, gde_13, n = 1000)
fire_count_14_sample <- stratified_sample(fire_count_14, gde_14, n = 1000)
fire_count_78_sample <- stratified_sample(fire_count_78, gde_78, n = 1000)
fire_count_80_sample <- stratified_sample(fire_count_80, gde_80, n = 1000)
fire_count_81_sample <- stratified_sample(fire_count_81, gde_81, n = 1000)
fire_count_85_sample <- stratified_sample(fire_count_85, gde_85, n = 1000)


# Prepare a list of the sample objects to input in the bootstrapping function. This allows us to calculate p-values for all objects at once. 
fire_count_sample_list <- list()

for(i in 1:13){
  obj <- (paste0('fire_count_', eco_region_codes[i], '_sample'))
  fire_count_sample_list <- c(fire_count_sample_list, obj)
}

```

### Bootstrapping
For this analysis, this code uses a difference in means bootstrapping approach. This means that the sample data created above is resampled 1000 times with replacement, and the difference in mean fire count in GDEs vs. non-GDEs (non-GDE mean minus GDE mean) is calculated for each resample. Then, we analyze the distribution of the 1000 difference in means and determine if the estimated population difference in means is significantly different from 0 or not. For example: If the difference in means is 0, that would mean that there is no difference in fire counts in GDEs when compared to non-GDEs in that ecoregion. 

```{r}
# Perform difference in means bootstrapping on each sample in our list of sample objects created above.  
fire_count_bootstrap <- bootstrap_results(fire_count_sample_list) 

# Bind with ecoregion names and geometries, and determine results from bootstrapping for each ecoregion. THIS IS NOT A SPATIAL OBJECT, see immediately below this object for a spatial object. 
fire_count_results <- fire_count_bootstrap %>%
  bind_cols(eco_regions_names) %>%
  inner_join(eco_regions_geoms, by = "us_l3code") %>% 
  mutate(result = case_when(
    p_value < 0.05 & confidence_interval_upper < 0 ~ "Reject H0 - Incr. Fire Count",
    p_value < 0.05 & confidence_interval_lower > 0 ~ "Reject H0 - Decr. Fire Count",
    p_value > 0.05 ~ "Cannot Reject H0")) %>%
  mutate(p_value = format(p_value, scientific = FALSE))

# Create a spatial object for mapping 
fire_count_results_sp <- fire_count_results %>%
  subset(select = c("p_value","result","region", "geometry")) %>%
  st_as_sf(crs = crs_ca$wkt)  
```

### Map Results of Difference in Means test
```{r}
# Light grey: #EDF0F0
# Red: #9D2A38
# Teal: #2A9D8F
# Darker grey: #C8D5D3
color_palette <- c("Reject H0 - Incr. Fire Count" = "#9D2A38",
                   "Reject H0 - Decr. Fire Count" = "#2A9D8F",
                   "Cannot Reject H0" = "#C8D5D3")


tm_shape(fire_count_results_sp) + 
  tm_polygons + 
  tm_fill(col = "result", palette = color_palette, title = "Result") + 
  tm_borders(col = "white", 
             lwd = 0.5) + 
  tm_layout(main.title = "",
            main.title.size = 0.8,
            frame = FALSE) 
```

# Calculate Summary Statistics 
This code chunk uses the fire_count_stats() function in the "Functions" folder to calculate some summary statistics for fire counts in each ecoregion. It calculates the mean, maximum, minimum, and mode (excluding 0s- burned cells only) fire count for the input sample data. 
```{r}
fire_count_1_stats <- fire_count_stats(fire_count_1_sample, eco_region_code = 1)
fire_count_4_stats <- fire_count_stats(fire_count_4_sample, eco_region_code = 4)
fire_count_5_stats <- fire_count_stats(fire_count_5_sample, eco_region_code = 5)
fire_count_6_stats <- fire_count_stats(fire_count_6_sample, eco_region_code = 6)
fire_count_7_stats <- fire_count_stats(fire_count_7_sample, eco_region_code = 7)
fire_count_8_stats <- fire_count_stats(fire_count_8_sample, eco_region_code = 8)
fire_count_9_stats <- fire_count_stats(fire_count_9_sample, eco_region_code = 9)
fire_count_13_stats <- fire_count_stats(fire_count_13_sample, eco_region_code = 13)
fire_count_14_stats <- fire_count_stats(fire_count_14_sample, eco_region_code = 14)
fire_count_78_stats <- fire_count_stats(fire_count_78_sample, eco_region_code = 78)
fire_count_80_stats <- fire_count_stats(fire_count_80_sample, eco_region_code = 80)
fire_count_81_stats <- fire_count_stats(fire_count_81_sample, eco_region_code = 81)
fire_count_85_stats <- fire_count_stats(fire_count_85_sample, eco_region_code = 85)

# Bind results into one dataframe 
fire_count_stats_all <- rbind(fire_count_1_stats, fire_count_4_stats, 
                              fire_count_5_stats, fire_count_6_stats,
                              fire_count_7_stats, fire_count_8_stats,
                              fire_count_9_stats, fire_count_13_stats, 
                              fire_count_14_stats, fire_count_78_stats,
                              fire_count_80_stats, fire_count_81_stats,
                              fire_count_85_stats)
```

### Visualizations
This code chunk creates a histogram showing the distribution of fire counts in GDEs and non-GDEs for each ecoregion in the sample list, created in the "Obtain Sample Data" section above. 
```{r}
# Loop through each value in the list and ecoregion code vector
for (i in seq_along(fire_count_sample_list)) {
  # Get the current object name and eco region code
  sample <- get(fire_count_sample_list[[i]])
  eco_region_code <- eco_region_codes[i]

  # Call the function with the object name and eco region code
  plot <- fire_count_histogram(sample, eco_region_code)

  # Save the plot as a png file, automatically named based on the ecoregion code. 
  ggsave(
    filename = paste0("fire_count_histogram_", eco_region_code, ".png"),
    plot = plot,
    device = "png",
    path = "plots/",
    units = c("in"),
    width = 11,
    height = 7
  )
dev.off()
}
```

